Automated reading of residual plots with computer vision models

### 2023-11-20 paper/paper.Rmd
update paper

---
title: |
  Automated reading of residual plots with computer vision models
type: ARTICLE TEMPLATE
author:
  - name: Weihao Li
    affil: a
    email: weihao.li@monash.edu
affiliation:
  - num: a
    address: |
      Department of Econometrics and Business Statistics, Monash University, Clayton, VIC, Australia
bibliography: ref.bib
abstract: |
  TBD.
keywords: |
  TBD
header-includes: |
  \usepackage{lscape}
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}
output: rticles::tf_article
editor_options: 
  chunk_output_type: console
---

@@

Delivering a residual plot as a lineup is generally considered a good practice. However, as pointed out by @li2023plot, a primary limitation of the lineup protocol is its reliance on human judgments. Unlike conventional statistical tests that can be performed numerically and automatically in statistical software, the lineup protocol requires human evaluation of images. This characteristic makes it less suitable for large-scale applications, given the associated high labour costs and time requirements.

In this paper, we attempt to address this issue by integrating computer vision models into residual plot diagnostics. The paper is structured as follows: ...

@@

<!-- # Related works? -->

<!-- # Literature review -->

<!-- ## Visual inference in residual diagnostics -->

<!-- ## Existing approaches to automated plot analysis -->

<!-- ### Computer vision approaches -->

<!-- ### Other approaches -->

>>>>

---
title: |
  Automated reading of residual plots with computer vision models
type: ARTICLE TEMPLATE
author:
  - name: Weihao Li
    affil: a
    email: weihao.li@monash.edu
affiliation:
  - num: a
    address: |
      Department of Econometrics and Business Statistics, Monash University, Clayton, VIC, Australia
bibliography: ref.bib
abstract: |
  TBD.
keywords: |
  TBD
header-includes: |
  \usepackage{lscape}
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}
  \doublespacing
output: rticles::tf_article
editor_options: 
  chunk_output_type: console
---

@@

The practice of delivering a residual plot as a lineup is generally regarded as a valuable approach. Beyond its application in residual diagnostics, the lineup protocol has  integrated into the analysis of diverse subjects. For instance, \cite{loy2013diagnostic, loy2014hlmdiag, loy2015you} illustrate its applicability in diagnosing hierarchical linear models. Additionally, @widen2016graphical demonstrates its utility in geographical research, while @krishnan2021hierarchical explores its effectiveness in forensic examinations.

However, as pointed out by @li2023plot, a primary limitation of the lineup protocol lies in its reliance on human judgments. Unlike conventional statistical tests that can be performed numerically and automatically in statistical software, the lineup protocol requires human evaluation of images. This characteristic makes it less suitable for large-scale applications, given the associated high labour costs and time requirements.

There is a compelling need to develop an approach that alleviates people's workload by automating repetitive tasks and providing standardized results in a controlled environment. The large-scale evaluation of lineups is impractical without the use of technology and machines.

The utilization of computers to interpret data plots has a rich history, with early efforts such as "Scagnostics" by @tukey1985computer, focusing on scatterplot diagnostics. @wilkinson2005graph expanded on this work in 2005, introducing graph theoretic scagnostics, which encompassed nine computable measures applied to planar proximity graphs. These measures, including "Outlying," "Skinny," "Stringy," "Straight," "Monotonic," "Skewed," "Clumpy," and "Striated," aimed to characterize outliers, shape, density, trend, and coherence of the data.

While this approach has been inspiring, there is a recognition that it may not capture all the necessary visual features distinguishing actual residual plots from null plots. An alternative, more promising approach involves allowing computers to learn the visual feature extraction function. In essence, this means enabling computers to discern which visual features are significant for residual diagnostics and how to extract them. Modern computer vision models, a subset of artificial intelligence, present a compelling solution to this challenge.

Modern computer vision models often rely on deep neural networks with convolutional layers [@fukushima1982neocognitron]. These layers leverage hierarchical patterns in data, downsizing and transforming images by summarizing information in a small space. Numerous studies have demonstrated the efficacy of convolutional layers in addressing various vision tasks, including image recognition [@rawat2017deep].

Despite the widespread use of computer vision models in fields like computer-aided diagnosis [@lee2015image], pedestrian detection [@brunetti2018computer], and facial recognition [@emami2012facial], their application in reading data plots remains limited. While some studies have explored the use of computer vision models for tasks such as reading recurrence plots for time series regression [@ojeda2020multivariate], time series classification [@chu2019automatic; @hailesilassie2019financial; @hatami2018classification; @zhang2020encoding], anomaly detection [@chen2020convolutional], and pairwise causality analysis [@singh2017deep], the application of reading residual plots with computer vision models represents a relatively new field of study.

In this paper, we develop computer vision models and integrate them into the residual plots diagnostics workflow, filling the gap of.... The paper is structured as follows: ...

@@

### 2023-11-20 paper/paper.tex
update paper

\usepackage{lscape}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\usepackage{setspace}

@@

Delivering a residual plot as a lineup is generally considered a good
practice. However, as pointed out by \citet{li2023plot}, a primary
limitation of the lineup protocol is its reliance on human judgments.
Unlike conventional statistical tests that can be performed numerically
and automatically in statistical software, the lineup protocol requires
human evaluation of images. This characteristic makes it less suitable
for large-scale applications, given the associated high labour costs and
time requirements.

In this paper, we attempt to address this issue by integrating computer
vision models into residual plot diagnostics. The paper is structured as
follows: \ldots{}

>>>>

\usepackage{lscape}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\usepackage{setspace}
\doublespacing

@@

The practice of delivering a residual plot as a lineup is generally
regarded as a valuable approach. Beyond its application in residual
diagnostics, the lineup protocol has integrated into the analysis of
diverse subjects. For instance,
\cite{loy2013diagnostic, loy2014hlmdiag, loy2015you} illustrate its
applicability in diagnosing hierarchical linear models. Additionally,
\citet{widen2016graphical} demonstrates its utility in geographical
research, while \citet{krishnan2021hierarchical} explores its
effectiveness in forensic examinations.

However, as pointed out by \citet{li2023plot}, a primary limitation of
the lineup protocol lies in its reliance on human judgments. Unlike
conventional statistical tests that can be performed numerically and
automatically in statistical software, the lineup protocol requires
human evaluation of images. This characteristic makes it less suitable
for large-scale applications, given the associated high labour costs and
time requirements.

There is a compelling need to develop an approach that alleviates
people's workload by automating repetitive tasks and providing
standardized results in a controlled environment. The large-scale
evaluation of lineups is impractical without the use of technology and
machines.

The utilization of computers to interpret data plots has a rich history,
with early efforts such as ``Scagnostics'' by \citet{tukey1985computer},
focusing on scatterplot diagnostics. \citet{wilkinson2005graph} expanded
on this work in 2005, introducing graph theoretic scagnostics, which
encompassed nine computable measures applied to planar proximity graphs.
These measures, including ``Outlying,'' ``Skinny,'' ``Stringy,''
``Straight,'' ``Monotonic,'' ``Skewed,'' ``Clumpy,'' and ``Striated,''
aimed to characterize outliers, shape, density, trend, and coherence of
the data.

While this approach has been inspiring, there is a recognition that it
may not capture all the necessary visual features distinguishing actual
residual plots from null plots. An alternative, more promising approach
involves allowing computers to learn the visual feature extraction
function. In essence, this means enabling computers to discern which
visual features are significant for residual diagnostics and how to
extract them. Modern computer vision models, a subset of artificial
intelligence, present a compelling solution to this challenge.

Modern computer vision models often rely on deep neural networks with
convolutional layers \citep{fukushima1982neocognitron}. These layers
leverage hierarchical patterns in data, downsizing and transforming
images by summarizing information in a small space. Numerous studies
have demonstrated the efficacy of convolutional layers in addressing
various vision tasks, including image recognition \citep{rawat2017deep}.

Despite the widespread use of computer vision models in fields like
computer-aided diagnosis \citep{lee2015image}, pedestrian detection
\citep{brunetti2018computer}, and facial recognition
\citep{emami2012facial}, their application in reading data plots remains
limited. While some studies have explored the use of computer vision
models for tasks such as reading recurrence plots for time series
regression \citep{ojeda2020multivariate}, time series classification
\citep{chu2019automatic, hailesilassie2019financial, hatami2018classification, zhang2020encoding},
anomaly detection \citep{chen2020convolutional}, and pairwise causality
analysis \citep{singh2017deep}, the application of reading residual
plots with computer vision models represents a relatively new field of
study.

In this paper, we develop computer vision models and integrate them into
the residual plots diagnostics workflow, filling the gap of\ldots. The
paper is structured as follows: \ldots{}
